Cyclone: High Availability for Persistent Key Value Stores

### 2017-11-19 paper/paper.tex
Updates.

\title{Cyclone: High Availability for Persistent Key Value Stores}
\author{Amitabha Roy \and Subramanya R. Dulloor}

@@

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many distributed data
serving solutions with innovations targeted at taking advantage of growing flash
speeds. Unfortunately their performance is hampered by the need to maintain and
replicate a write ahead log to guarantee availability in the face of machine and
storage failures. Cyclone is a replicated log plug-in for key value stores that
systematically addresses various sources of this bottleneck. It uses a small
amount of non-volatile memory directly addressable by the CPU - such as in the
form of NVDIMMs or Intel 3DXPoint - to remove block oriented IO devices such as
SSDs from the critical path for appending to the log. This enables it to address
network overheads using an implementation of the RAFT consensus protocol that is
designed around a userspace network stack to relieve the CPU of the burden of
data copies. Finally, it provides a way to efficiently map the commutativity in
key-value store APIs to the parallelism available in commodity NICs.  Cyclone is
able to replicate millions of small updates per second using only commodity 10
gigabit ethernet adapters. As a practical application, we use it to improve the
performance (and availability) of RocksDB, a popular persistent key value store
by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software, once
block IO is removed from the critical path. Existing work on log replication
using state machine replication protocols such as Paxos~\cite{paxos} and
RAFT~\cite{raft} has not - thus far - addressed high performance replication in
the local area network of a datacenter using commodity networks.  Research
focusing on addressing wide area network latencies by reducing the number of
network hops with new protocols such as Fast Paxos~\cite{fast-paxos} or network
switch modifications such as No Paxos~\cite{nopaxos} do not address throughput
bottlenecks due to network stack related inefficiencies in the participating
nodes of the quorum -- a more urgent problem with local area replication. On the
other hand, work done to address network related inefficiencies start with the
assumption that the bottleneck to good throughput on the network is the CPU and
therefore one must either resort to network offload mechanisms such as RDMA~\cite{dare,
  farm, faast} or offload the protocol entirely to an
FPGA~\cite{consensus_box}. We demonstrate in this paper that this is not the
case. Rather, once persistence is provided by directly attached non-volatile
memory, thereby eliminating block IO from the persistence step in replication
protocols, the problem becomes akin to multicasting the same log entry
from the leader to all followers. We design Cyclone's network stack around well
known principles in the software packet switching community. In particular, a
careful implementation of a consensus protocol that relieves the CPU of the
responsibility of data movement permits a high performance implementation
entirely in software that addresses the order of magnitude network related
performance gap in Figure~\ref{fig:problem} using only commodity 10 Gigabit
ethernet.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. The third
key contribution of this work is to show how one can exploit this parallelism in
Cyclone by making use of the commutativity inherent in key-value store
interfaces and mapping it to the available parallelism in the system via
multiple instances of the same consensus protocol.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients.  Cyclone guarantees that the order of operations to any key are linearizable. 
A key sees the exact same sequence of update operations (including deletes) 
on all replicas. These per-key guarantees follow immediately from the 
properties of the RAFT log and by requiring clients to perform a quorum read (logged
as an entry in the RAFT log).

In addition, Cyclone also provides clients with a weak read that does not require a 
quorum read thereby requiring only one round trip to a server.  A weak read of a key sees a prefix of the linearizable history of a key that 
includes all successful update operations made by the client before it issues the 
weak read. To do a weak read, 
clients maintain the last known term of the RAFT leader locally and ensure they always talk
to a RAFT leader that has at least the same term. This ensures that clients see their own writes
avoiding split brain problems where a network partitions allows clients to
talk to old RAFT leaders after having committed changes to a new RAFT leader. In addition, a weak read waits until all pending RAFT log entries have been committed and applied before executing. This ensures that in the event of a failover, any previous committed updates from the same client have been applied and are visible to the weak read. A weak read is not linearizable as the RAFT leader responding to the weak read might actually no longer be the leader - a new one having been elected 
without it realizing, perhaps due to a network partition. This means that while a weak read from a client sees its own writes in this (perhaps pathological) case it might not see later writes by other clients.

Finally, Cyclone allows operations to multiple keys in an atomic batch, where
either all updates in the batch are applied or no operation in the batch is
applied - a property that holds \emph{at all replicas}, regardless of failures. We also use a lightweight batched operation to support snapshots. We discuss how batched operations are implemented later in the paper.

@@

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU, rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. However, directly attached persistent 
memory in the form of NVDIMMs today or new memory technologies in the
future~\cite{pmfs} will be of lesser capacity than traditional (NAND based)
flash available through an
SSD. Providing the same capacity as existing flash based logs for key value
stores therefore requires us to provide a second level of the log
placed on a flash SSD (called flashlog).  Entries are drained from
the NVM log to this flashlog in conveniently large units that are a multiple of
4KB -- the optimum IO unit for flash SSDs. This ensures we get the best possible
throughput from a flash device without paying the price of synchronous IO for
small key-value pairs. At the same time, the amount of non-volatile memory
required is small enough to not add undue cost to the server.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.4]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

@@

\begin{figure*}
  \begin{minipage}{.2\textwidth}
    \scriptsize
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Event handling}
\label{fig:control_plane}
\end{minipage}
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{minipage}
\begin{minipage}{.3\textwidth}
\centering
\includegraphics[scale=0.3]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{minipage}
\end{figure*}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

Although the work described in the previous section significantly boosts the
network performance of replication in Cyclone, it still does not come close to
saturating the capabilities of even the commodity NICs that we use. This is
because NICs today encapsulate significant parallelism in terms of
multiple send and receive queue pairs. Exploiting this capability to improve
throughput requires us to remove the bottleneck of a single sequential log in
RAFT. To this end, we extend the implementation described thus far to run
multiple copies of the RAFT consensus protocol each maintaining and replicating
its own two level log. We refer to these logs as physical logs. All these
instances however exist in the same process address space as the key value store
application itself and therefore manipulate the same application.

@@

\subsection{Ganged Operations}
Supporting batched updates to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together even in the face of failures. One option is to have an
external co-ordinator run a distributed transaction using two phase commit. This
requires extra external resources to host the co-ordinator what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. This also 
requires two round trips to the
leader to complete  We therefore describe a simpler and more performant
solution called ``ganged operations" in Cyclone. Ganged operations allow a
stateless client to issue a batch of update operations as a ``gang",
simultaneously and atomically updating a batch of keys in a single call. This preserves
the system architecture of existing deployments of single image key value store such as Rocksdb.

The first key problem we need to deal with is that the leaders for the different
RAFT instances to which the keys map can be located on different machines. To
avoid having a client co-ordinate different machines for a single operation, we
constrain the RAFT leader election algorithm for the different RAFT instances to
converge to the same machine. We do this by triggering re-elections for a RAFT
instance other than that of the first physical log until it is on the same
machine as the \emph{first} physical log. This process is therefore resilient to
failed machines. RAFT requires that the leader have the most up to date log. To
ensure that the process converges we therefore ensure that the currently elected
leader brings a majority quorum up to date before triggering a reelection. We
also do not accept any requests from clients until all leaders are
co-located. We assume that no network partitions occur \emph{within a machine}
i.e. some RAFT instances are able to send and receive packets, while others are
not, a possibility we discount due to our single process design.

@@

\begin{figure}
  \centering
  \scriptsize
  \bf
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrier.mask != mask of all participants
        if barrier.failed
         send retry to client
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         send retry to client
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure}

@@

\begin{figure*}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/replicas.pdf}
  \caption{Impact of replica count}
  \label{fig:replicas}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/512.pdf}
  \caption{Impact of payload size}
  \label{fig:payload}
\end{minipage}
\begin{minipage}{.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog.pdf}
  \caption{Adding second level log}
  \label{fig:flashlog}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{minipage}
\end{figure*}

@@

\begin{figure*}
  \begin{minipage}{0.33\textwidth}
    \centering
    \includegraphics[scale=0.45]{results2/rocksdb.pdf}
    \caption{Rocksdb - small updates}
    \label{fig:rocksdb}
  \end{minipage} 
  \begin{minipage}{0.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
  \end{minipage} 
  \begin{minipage}{0.33\textwidth}
    \includegraphics[scale=0.45]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
  \end{minipage}
\end{figure*}

@@

Our evaluation has, thus far, focused on synthetic write heavy workloads. To
illustrate that Cyclone can also be useful for real-world read-heavy workloads, 
we configured our clients to generate key value requests that mimic the
distribution of value sizes in Facebook's ETCD
trace~\cite{fb_workload_analysis,fb_scaling_memcache} and are read heavy (95\%
reads). Figure~\ref{fig:fb} illustrates that even with a read dominated
workload, Cyclone is effective at providing a write ahead log with 3 way
replication with a lower performance overhead than Rocksdb's own write ahead
logging. We consider the case of another real-world workload that is more write 
intensive (at 80\%) and with a smaller range of value sizes -- derived from
Facebook's VAR trace~~\cite{fb_workload_analysis,fb_scaling_memcache}, in
Figure~\ref{fig:fb2}. One can similarly conclude that Cyclone provides more
functionality (replicated write ahead log) at a far lower performance penalty
that Rocksdb's existing single machine write ahead log solution.

\begin{figure*}
\begin{minipage}{0.33\textwidth}
\includegraphics[scale=0.45]{results2/fb.pdf}
\caption{Rocksdb - read heavy}
\label{fig:fb}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/fb2.pdf}
  \caption{Rocksdb - write heavy}
  \label{fig:fb2}
\end{minipage}
\begin{minipage}{0.33\textwidth}
  \includegraphics[scale=0.45]{results2/failover.pdf}
  \caption{Rocksdb - failover}
  \label{fig:timeline}
\end{minipage}
\end{figure*}

@@

%The scalable commutativity rule~\cite{scalable_commutativity} states that for a
%shared memory program to scale in performance with concurrency it must also have
%an interface that is commutative for common case invocations. We have also used
%interface level commutativity to extract parallelism for the purpose of scaling
%replication performance. Others have used the same commutativity to create
%per-cpu data structures to scale KV store performance~\cite{flodb}.

>>>>

\title{Cyclone: High Availability for Persistent Key Value Stores}
\author{Amitabha Roy \\ amitabha.roy@gmail.com 
	     \and Subramanya R. Dulloor \\ dulloor@gmail.com}

@@

\begin{abstract}
\vspace{0.05in}
Persistent key value stores are an important component of many distributed data
serving solutions with innovations targeted at taking advantage of growing flash
speeds. Unfortunately their performance is hampered by the need to maintain and
replicate a write ahead log to guarantee availability in the face of machine and
storage failures. Cyclone is a replicated log plug-in for key value stores that
systematically addresses various sources of this bottleneck. It uses a small
amount of non-volatile memory directly addressable by the CPU - such as in the
form of NVDIMMs or Intel 3DXPoint - to remove block oriented IO devices such as
SSDs from the critical path for appending to the log. This enables it to address
network overheads using an implementation of the RAFT consensus protocol that is
designed around a userspace network stack to relieve the CPU of the burden of
data copies.  Finally, it provides a way to exploit the parallelism available in commodity NICs.  
Cyclone is able to replicate millions of small updates per second using only commodity 10
gigabit ethernet adapters. As a practical application, we use it to improve the
performance (and availability) of RocksDB, a popular persistent key value store
by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

The second key contribution of this work is in showing that the throughput
bottlenecks in the network stack can be mitigated entirely in software, once
block IO is removed from the critical path. Existing work on log replication
using state machine replication protocols such as Paxos~\cite{paxos} and
RAFT~\cite{raft} has not - thus far - addressed high performance replication in
the local area network of a datacenter using commodity networks.  Research
focusing on addressing network latencies by reducing the number of
network hops with new protocols such as Fast Paxos~\cite{fast-paxos} or network
switch modifications such as No Paxos~\cite{nopaxos} do not address throughput
bottlenecks due to network stack related inefficiencies in the participating
nodes of the quorum -- a more urgent problem with local area replication. On the
other hand, work done to address network related inefficiencies start with the
assumption that the bottleneck to good throughput on the network is the CPU and
therefore one must either resort to network offload mechanisms such as RDMA~\cite{dare,
  farm, faast} or offload the protocol entirely to an
FPGA~\cite{consensus_box}. We demonstrate in this paper that this is not the
case. Rather, once persistence is provided by directly attached non-volatile
memory, thereby eliminating block IO from the persistence step in replication
protocols, the problem becomes akin to multicasting the same log entry
from the leader to all followers. We design Cyclone's network stack around well
known principles in the software packet switching community. In particular, a
careful implementation of a consensus protocol that relieves the CPU of the
responsibility of data movement permits a high performance implementation
entirely in software that addresses the order of magnitude network related
performance gap in Figure~\ref{fig:problem} using only commodity 10 Gigabit
ethernet.

Both the storage (NVM, SSD) as well as the NIC expose significant parallelism
that is left unused when Cyclone is used to replicate a single log. The third
key contribution of this work is to show how one can exploit this parallelism in
Cyclone automatically creates and manages multiple logs, mapping them to the 
available parallelism in the system via multiple instances of the same consensus 
protocol.

@@

Cyclone uses the RAFT consensus protocol~\cite{raft} to keep the logs in sync
across the replicas. On failure of the leader replica, a new leader is
automatically elected -- RAFT ensures that the new leader has the most up to
date log. The Cyclone client library automatically locates the new leader on a
failover, resulting in only a brief interruption in service for
clients.  Cyclone guarantees that the order of operations to any particular key 
are linearizable. A key sees the exact same sequence of update operations 
(including deletes) on all replicas. These per-key guarantees follow immediately 
from the properties of the RAFT log and by requiring clients to perform a quorum 
read (logged as an entry in the RAFT log).

In addition, Cyclone also provides clients with a weak read that does not require a 
quorum read thereby requiring only one round trip to a server.  A weak read of a key sees a 
prefix of the linearized sequence of updates to a key that includes all successful update operations made 
by the client before it issues the weak read. To do a weak read, 
clients maintain the last known term of the RAFT leader locally and ensure they always talk
to a RAFT leader that has at least the same term. This ensures that clients see their own writes
avoiding split brain problems where a network partition allows clients to
talk to old RAFT leaders after having committed changes to a new RAFT leader. In addition, a weak 
read waits until all pending RAFT log entries have been committed and applied before executing. 
This ensures that in the event of a failover, any previous committed updates from the same client 
have been applied and are visible to the weak read. A weak read is not linearizable as the RAFT 
leader responding to the weak read might actually no longer be the leader - a new one having been 
elected without it realizing, perhaps due to a network partition. This means that while a weak read 
from a client sees its own writes, in some rare cases it might not see later writes by other clients.

Finally, Cyclone allows operations to multiple keys in an atomic batch, where
either all updates in the batch are applied or no operation in the batch is
applied - a property that holds \emph{at all replicas}, regardless of failures. Batched operations 
are supported in Rocksdb to allow referential integrity between keys - where the value associated
with a  key is a reference to another key. We also use a lightweight batched operation to support 
snapshots - that can be viewed as an atomic no-op to all keys. We discuss how batched operations are 
supported by Cyclone later in the paper.

@@

Cyclone adopts a different strategy by making use of novel memory technologies
to obviate the need to make this tradeoff in the first place. We assume a small
amount of non-volatile memory directly addressable by the CPU, rather than being
placed behind a block oriented IO interface. Log appends are done to the ``NVM
log'' placed in this non-volatile memory. However, directly attached persistent 
memory in the form of NVDIMMs today or new memory technologies in the
future~\cite{pmfs} will be of lesser capacity than traditional (NAND based)
flash available through an
SSD. Providing the same capacity as existing flash based logs for key value
stores therefore requires us to provide a second level of the log
placed on a flash SSD (called flashlog).  Entries are drained from
the NVM log to this flashlog in conveniently large units that are a multiple of
4KB -- the optimum IO unit for flash SSDs. This ensures we get the best possible
throughput from a flash device without paying the price of synchronous IO for
small key-value pairs. At the same time, the amount of non-volatile memory
required is small enough to not add undue cost to the server. This is in contrast to
systems like FARM~\cite{farm} that require both the key value store and write ahead log
to be held entirely in non-volatile memory.

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.6]{figures2/flashlog_page.pdf}
  \caption{Flashlog segment}
  \label{fig:flashlog_page}
\end{figure}

@@

\begin{figure}
\small
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Event handling}
\label{fig:control_plane}
\end{figure}

\begin{figure*}
  \centering
  \includegraphics[scale=0.6]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure*}

\begin{figure}
\centering
\includegraphics[scale=0.5]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[scale=0.5]{figures2/batching.pdf}
  \caption{Batching}
  \label{fig:batching}
\end{figure}

@@

Although the work described in the previous section significantly boosts the
network performance of replication in Cyclone, it still does not come close to
saturating the capabilities of even the commodity NICs that we use. This is
because NICs today encapsulate significant parallelism in terms of
multiple send and receive queue pairs. Exploiting this capability to improve
throughput requires us to remove the bottleneck of a single sequential log in
RAFT. To this end, we extend the implementation described thus far to run
multiple copies of the RAFT consensus protocol each maintaining and replicating
its own two level log. We refer to these logs as physical logs. All these
instances however exist in the same process address space as the key value store
application itself and therefore manipulate the same application. The number of 
instances is a fixed property of the Cyclone service and cannot be changed after 
startup.

@@

\subsection{Ganged Operations}
Using multiple instances of RAFT leads to a serious problem with batched operations, 
since we need to split up the batch into a mini-batch for each of the physical logs. There
is however no guarantee that the request for a mini-batch will succeed breaking the atomicity
requirement for batched updates. The obvious and simple solution is to do two phase commit across
the logs, ensuring all or nothing semantics. However this was unacceptable to some customers who
pointed out that ensuring that an external co-ordinator implementing two phase commit does not itself
fail requires running a separate replicated service - with its attendant resource and reliability headaches.
We were therefore challenged to come up with a solution that did not require two phase commit.

Our final solution is based on two observations. First, we \emph {do not} require that either all mini-batches
icommit to their raft logs or none do, What we actually requires is that either all are \emph{applied} to Rocksdb or 
none of them are. Put another way we should not apply a mini-batch from a ganged operation if any of the other mini-batches
have failed to commit to their RAFT log. This means that the client can be stateless. The second observation is that
checking whether the other minibatches have committed is easy because the different RAFT instances can communicate 
through shared memory on the same machines unlike the more general case of distributed participants in two phase 
commit.

We now describe our solution called ``ganged operations" in Cyclone. 
The first key problem we need to deal with is that the leaders for the different
RAFT instances to which the keys map can be located on different machines. To
avoid having a client co-ordinate different machines for a single operation, we
constrain the RAFT leader election algorithm for the different RAFT instances to
converge to the same machine. We do this by triggering re-elections for a RAFT
instance other than that of the first physical log until it is on the same
machine as the \emph{first} physical log. This process is therefore resilient to
failed machines. RAFT requires that the leader have the most up to date log. To
ensure that the process converges we therefore ensure that the currently elected
leader brings a majority quorum up to date before triggering a reelection. We
also do not accept any requests from clients until all leaders are
co-located. We assume that no network partitions occur \emph{within a machine}
i.e. some RAFT instances are able to send and receive packets, while others are
not, a possibility we discount due to our single process design.

@@

\begin{figure*}
  \centering
  \small
  \bf
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrier.mask != mask of all participants
        if barrier.failed
         send retry to client
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         send retry to client
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

@@

\begin{figure}
  \includegraphics[scale=0.6]{results2/replicas.pdf}
  \caption{Impact of replica count}
  \label{fig:replicas}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/512.pdf}
  \caption{Impact of payload size}
  \label{fig:payload}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/flashlog.pdf}
  \caption{Adding second level log}
  \label{fig:flashlog}
\end{figure}

@@

\begin{figure}
  \includegraphics[scale=0.6]{results2/flashlog_512.pdf}
  \caption{Second level log - 512 bytes payload}
  \label{fig:flashlog_512}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/multi.pdf}
  \caption{Ganged Operations}
  \label{fig:ganged}
\end{figure}

\begin{figure}
  \includegraphics[scale=0.6]{results2/kv_baseline.pdf}
  \caption{Baseline KV replication}
  \label{fig:kv_baseline}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[scale=0.6]{results2/rocksdb.pdf}
    \caption{Rocksdb - small updates}
    \label{fig:rocksdb}
  \end{figure}

  \begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_256.pdf}
    \caption{Rocksdb - large updates}
    \label{fig:rocksdb_256}
  \end{figure}

  \begin{figure}
    \includegraphics[scale=0.6]{results2/rocksdb_multi.pdf}
    \caption{Rocksdb - batched writes}
    \label{fig:rocksdb_multi}
\end{figure}

@@

Our evaluation has, thus far, focused on synthetic write heavy workloads. To
illustrate that Cyclone can also be useful for real-world read-heavy workloads, 
we configured our clients to generate key value requests that mimic the
distribution of value sizes in Facebook's ETCD
trace~\cite{fb_workload_analysis,fb_scaling_memcache} and are read heavy (95\%
reads). We use weak reads in Cyclone to avoid the extra round trip for quorum reads.
Figure~\ref{fig:fb} illustrates that even with a read dominated
workload, Cyclone is effective at providing a write ahead log with 3 way
replication with a lower performance overhead than Rocksdb's own write ahead
logging. We consider the case of another real-world workload that is more write 
intensive (at 80\%) and with a smaller range of value sizes -- derived from
Facebook's VAR trace~~\cite{fb_workload_analysis,fb_scaling_memcache}, in
Figure~\ref{fig:fb2}. One can similarly conclude that Cyclone provides more
functionality (replicated write ahead log) at a far lower performance penalty
that Rocksdb's existing single machine write ahead log solution.

\begin{figure}
\includegraphics[scale=0.6]{results2/fb.pdf}
\caption{Rocksdb - read heavy}
\label{fig:fb}
\end{figure}
\begin{figure}
  \includegraphics[scale=0.6]{results2/fb2.pdf}
  \caption{Rocksdb - write heavy}
  \label{fig:fb2}
\end{figure}
\begin{figure}
  \includegraphics[scale=0.6]{results2/failover.pdf}
  \caption{Rocksdb - failover}
  \label{fig:timeline}
\end{figure}

@@

%The scalable commutativity rule~\cite{scalable_commutativity} states that for a
%shared memory program to scale in performance with concurrency it must also have
%an interface that is  for common case invocations. We have also used
%interface level commutativity to extract parallelism for the purpose of scaling
%replication performance. Others have used the same commutativity to create
%per-cpu data structures to scale KV store performance~\cite{flodb}.
